The Legacy of the European Post-Master in Urbanism at TU Delft: A Text Mining Approach

### 2023-01-12 analysis/paper/paper.qmd
Update setup code in paper

# Load packages ----
library(here)
library(dplyr)
library(readr)
library(stringr)
library(tidytext)
library(pluralize)
library(tm)
devtools::load_all(".")

# Read thesis metadata ----
emu_theses <- read_csv(here::here("analysis","data", "raw_data", "theses-metadata.csv")) %>%
  mutate(text = "")

pdf_names <- dir(here::here("analysis", "data", "raw_data"), pattern = "*.pdf")

emu_theses <- emu_theses %>%
  filter(file_name %in% pdf_names)

pdf_paths <- here::here("analysis","data", "raw_data", emu_theses$file_name)

emu_theses$text <- convert_pdf_text(pdf_paths)

@@

Stop words:
- Custom stop words from titles and author names should only be removed from corresponding theses;
- Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words;
- All theses were written in English; for common English stop words the "snowball" lexicon was used.

```{r}
stop_words_custom <- emu_theses %>% 
  unnest_tokens(output = word, input = title) %>%   ### add words from the title
  select(word) %>% 
  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis"))) %>% 
  rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
  rbind(., data.frame(word = c("hab", "km")))

@@

```{r}
# find the most frequently used words
emu_theses_top <- emu_theses_words %>%
  group_by(title) %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
  filter(n > 50) %>%              ### filter words with a minimum count
  ungroup() %>% 
  mutate(word = reorder(word, n))

emu_theses_top[emu_theses_top$word == "ring",]

emu_theses_top %>%
  ggplot(aes(x = word, y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2)

>>>>

# Load packages ----
## Package for managing paths
library(here)

## Metapackage for data science
library(tidyverse)

## Packages for text processing and analysis
library(tidytext)
library(SnowballC)
library(tm)

# Load analysis functions ----
devtools::load_all(".")

# Read thesis metadata ----
data_path <- here("analysis", "data", "raw_data")
pdf_names <- dir(path, pattern = "*.pdf")

emu_theses <- 
  read_csv(here(data_path, "theses-metadata.csv")) %>%
  mutate(text = "") %>%
  filter(!is.na(pdf_via))

pdf_paths <- here(data_path, emu_theses$file_name)
```

```{r}
#| label: extract-pdf-text
emu_theses$text <- convert_pdf_text(pdf_paths).
```

```{r}
#| label: clean-text

@@

Word-level tokenisation was applied with the `unnest_tokens()` function of the 
`tidytext` package. This was preferred over other tokenisers as it produced 
results that work seamlessly with other `tidyverse` tools.

Stop words:
- Three types of stop words: global, subject-specific and document stop words.
- Global stop words:
  - All theses were written in English, so for common English stop words the "snowball" lexicon was used
  - Optionally, try creating a custom stop word list on the entire corpus, by removing high-frequency words, starting with 20 words and increasing by 10 until reaching words that are not appropriate as stop words
- Subject-specific stop words:
  - Manually constructed list, needs domain knowledge
  - Words that are commonly used to structure a theses, such as "preface", "contents", and "introduction", were added as custom stop words
- Document stop words:
  - Stop words from titles and author names should only be removed from corresponding theses

```{r}
#| label: stop-words

stop_words_custom <- emu_theses %>% 
  unnest_tokens(output = word, input = title) %>%   ### add words from the title
  select(word) %>% 
  rbind(., data.frame(word = c("preface", "foreword", "introduction", "conclusion", "thesis", 
                               "source", "author"))) %>% 
  # rbind(., data.frame(word = c("city", "urban", "urbanism"))) %>% 
  rbind(., data.frame(word = c("hab", "km")))

@@

Stemming is a pre-processing step that needs to be carefully thought through, as 
it might either degrade topic modeling or produce no meaningful results [@hvitfeldt2021a].

```{r}
#| label: stemming

emu_theses_words %>% 
  mutate(stem = wordStem(words = word, 
                         language = "porter")) %>%  # Implement the Porter stemming algorithm for English provided by the SnowballC package
  count(stem, sort = TRUE)
```

```{r}
# Which are the top 20 most frequently used words in each thesis?
emu_theses_top_20 <- emu_theses_words %>%
  group_by(title) %>%
  count(word, sort = TRUE) %>%
  slice_max(n, n = 20) %>%        ### top_n is superseded; better use slice_max instead
  # filter(n > 50) %>%              ### filter words with a minimum count
  ungroup() %>% 
  mutate(word = reorder(word, n))

# Which are first 5 most frequently used words in each thesis?
emu_theses_first_5 <- emu_theses_words %>%
  count(title, word) %>% 
  group_by(title) %>% 
  arrange(desc(n)) %>% 
  slice(1:5)

# Which theses contain "fabric" in the top 20 most frequently used words?
emu_theses_top_20 %>% 
  filter(word == "fabric")

emu_theses_top_20 %>%
  slice(1:40) %>% 
  ggplot(aes(x = reorder(word, n), y = n, fill = title)) +
  geom_col() +
  xlab(NULL) +
  coord_flip() +
  labs(y = "Count",
       x = "Unique words",
       title = "Count of unique words found in thesis",
       subtitle = "Stop words were removed from the list") +
  facet_wrap( ~ title, scales = "free", ncol = 2)
